---
title: "Phago Assay Statistical Analysis"
author: "Brian"
date: "2022-10-18"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 2
---

<style type="text/css">
.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r include = F}
## Packages
if(!any(rownames(installed.packages()) == "pacman")){
   install.packages("pacman")
}

pacman::p_load(here,
               tidyverse,
               multcomp,
               multcompView,
               DT,
               forcats,
               ggtext,
               pals,
               emmeans,
               kableExtra,
               broom,
               insight,
               ggpubr,
               stringr,
               DescTools,
               forcats)

## Load RData
load(here("rdata","phago_tidy.rdata"))


```


```{r include= F}
## As Factor and reoder, and assign name rename
phago.assay$Treatment <- factor(phago.assay$Treatment, levels = c("no_treat", "AbO", "Chol", "Fru","LPS", "Chol_AbO","AbO_Fru", "AbO_LPS", "Chol_LPS", "Chol_Fru","Fru_LPS", "Chol_AbO_LPS","AbO_Fru_Chol_LPS"))

## New Dataframe reordered
phago.assay_reorder <- phago.assay[order(phago.assay$Treatment), ]

## Assign Names to Levels
levels(phago.assay_reorder$Treatment) <- list(Control  = "no_treat", 
                                                     AβO = "AbO",
                                                     Cholesterol = "Chol",
                                                     Fructose = "Fru",
                                                     LPS = "LPS",
                                                     AβO_Cholesterol = "Chol_AbO",
                                                     AβO_Fructose = "AbO_Fru",
                                                     AβO_LPS = "AbO_LPS",
                                                     Cholesterol_LPS = "Chol_LPS",
                                                     Fructose_Cholesterol = "Chol_Fru",
                                                     Fructose_LPS = "Fru_LPS",
                                                     AβO_Cholesterol_LPS = "Chol_AbO_LPS",
                                                     AβO_Fru_Chol_LPS = "AbO_Fru_Chol_LPS")

```

```{r include = F}
#Filter by sets
phago.assay_reorder_1 <- phago.assay_reorder %>%
        filter(Set %in% "1") %>%
        mutate(log2_fluorescence = log2(`Raw Recorded Fluorescence`))

phago.assay_reorder_2 <- phago.assay_reorder %>%
        filter(Set %in% "2") %>%
        mutate(log2_fluorescence = log2(`Raw Recorded Fluorescence`))

phago.assay_reorder_3 <- phago.assay_reorder %>%
        filter(Set %in% "3") %>%
        mutate(log2_fluorescence = log2(`Raw Recorded Fluorescence`))


```



## Phagocytosis Experiment SET 1 {.tabset .tabset-fade .tabset-pills}


### Figure (non-normalize to protein)

```{r echo = F, warning = F, fig.height = 10, fig.width = 12}
# ANOVA
anova <- aov(log2_fluorescence ~ Treatment, data = phago.assay_reorder_1)

# Extract P-value
p_value <- signif(summary(anova)[[1]][["Pr(>F)"]][1],3)

# Contrast with Dunnett
emm_contrasts <- anova %>%
  emmeans("Treatment") %>%
  contrast("trt.vs.ctrl", infer = c(TRUE, TRUE)) %>%
  as_tibble() %>%
  mutate(p.value.round = insight::format_p(p.value, name = NULL,),
         p.value.star = insight::format_p(p.value, name = NULL, stars_only = TRUE)) 

emm_contrasts_split <- emm_contrasts
## Split string
emm_contrasts_split[c('group1', 'group2')] <- str_split_fixed(emm_contrasts_split$contrast, ' - ', 2) 




## Data Summary ##
dt_summary <- phago.assay_reorder_1 %>%
  group_by(Treatment) %>%
        summarize(mean = mean(log2_fluorescence),
                  sd = sd(log2_fluorescence))
#define custom color scale
myColors <- as.vector(kelly(13))
names(myColors) <- levels(dt_summary$Treatment)
custom_colors <- scale_fill_manual(name = "Treatment", values = myColors)


################################## Create Plot ##################################
phago_figure <- dt_summary %>%
        mutate(Treatment = fct_reorder(Treatment, mean, .desc = FALSE)) %>%
ggplot() + 
        
        geom_bar(aes(y = mean, x = Treatment, fill = Treatment), stat = "identity", show.legend = TRUE) +
        
        geom_errorbar(data = dt_summary,
                aes(
                  ymin = mean - sd,
                  ymax = mean + sd,
                  x = Treatment
                ),
                width = 0.1) +
        
# Color Bar Plot (Discrete)
custom_colors + 
        
        labs(x = "Treatment", y = expression(paste("Phagocytosis, ",Log[2],"(Fluorescence)")), subtitle = paste("ANOVA:", " P"," = ", paste0(p_value, collapse = NULL))) +
        
# letters
stat_pvalue_manual(data = emm_contrasts_split,
                   y.position = rep(15,12),
                   label = "p.value.star",
                   x = "group1",
                   vjust = 2,
                   remove.bracket=TRUE,
                   size = 10) +
        theme_classic() +
        theme(axis.text = element_text(angle = 90, color = "black"),
              text = element_text(size = 23)) 
#phago_figure
#saveRDS(object = phago_figure, file = here("manuscript","data","phago.RDS"))
```


```{r include = F}
#ggsave(plot = phago_figure,here("manuscript","figures","phago_assay.jpg"), dpi = 300, width = 10, height = 8)

```



### Stats - Dunnett's test (non-normalize to protein)

```{r echo = F}
datatable(as.data.frame(emm_contrasts), colnames = c("Contrast", "Mean Diff", "SE", "DF"," Lower CI 95%", "Upper CI 95%","T.ratio","P.adjusted","P.rounded","Sig Level"))

datatable(dt_summary, colnames = c("Treatment","Mean","Standard Deviation (SD)")) %>%
        formatRound(columns = c("mean","sd"), digits = 2)
```



## Phagocytosis Experiment SET 2 - Manuscript {.tabset .tabset-fade .tabset-pills}


### Figure (non-normalize to protein)

```{r echo = F, warning = F, fig.height = 10, fig.width = 12}
# ANOVA
anova <- aov(log2_fluorescence ~ Treatment, data = phago.assay_reorder_2)

# Extract P-value
p_value <- signif(summary(anova)[[1]][["Pr(>F)"]][1],3)

# Contrast with Dunnett
emm_contrasts <- anova %>%
  emmeans("Treatment") %>%
  contrast("trt.vs.ctrl", infer = c(TRUE, TRUE)) %>%
  as_tibble() %>%
  mutate(p.value.round = insight::format_p(p.value, name = NULL,),
         p.value.star = insight::format_p(p.value, name = NULL, stars_only = TRUE)) 

emm_contrasts_split <- emm_contrasts
## Split string
emm_contrasts_split[c('group1', 'group2')] <- str_split_fixed(emm_contrasts_split$contrast, ' - ', 2) 




## Data Summary ##
dt_summary <- phago.assay_reorder_2 %>%
  group_by(Treatment) %>%
        summarize(mean = mean(log2_fluorescence),
                  sd = sd(log2_fluorescence))
#define custom color scale
myColors <- as.vector(kelly(13))
names(myColors) <- levels(dt_summary$Treatment)
custom_colors <- scale_fill_manual(name = "Treatment", values = myColors)

## Custom Factor where Control is First
#dt_summary$Treatment <- factor(dt_summary$Treatment, 
                              # levels = c("Control", 
                                         # "Cholesterol_LPS",
                                          #"Fructose_Cholesterol",
                                          #"AβO_Cholesterol_LPS",
                                          #"Fructose",
                                          #"AβO_Cholesterol",
                                         # "Cholesterol",
                                          #"LPS",
                                          #"Fructose_LPS",
                                          #"AβO_Fru_Chol_LPS",
                                         # "AβO",
                                         # "AβO_LPS",
                                         # "AβO_Fructose"))

################################## Create Plot ##################################
phago_figure <- dt_summary %>%
ggplot() + 
        geom_bar(aes(y = mean, x = Treatment, fill = Treatment), stat = "identity", show.legend = TRUE) +
        
        geom_errorbar(data = dt_summary,
                aes(
                  ymin = mean - sd,
                  ymax = mean + sd,
                  x = Treatment
                ),
                width = 0.1) +
        
# Color Bar Plot (Discrete)
custom_colors + 
        
        labs(x = "Treatment", y = expression(paste("Phagocytosis, ",Log[2],"(Fluorescence)")), subtitle =  expression(paste("ANOVA: P = ", "3.14 x 10"^-9))) +
        
# letters
stat_pvalue_manual(data = emm_contrasts_split,
                   y.position = rep(15,12),
                   label = "p.value.star",
                   x = "group1",
                   vjust = 2,
                   remove.bracket=TRUE,
                   size = 10) +
        theme_classic() +
        theme(axis.text = element_text(angle = 90, color = "black"),
              text = element_text(size = 23)) 
phago_figure
saveRDS(object = phago_figure, file = here("manuscript","data","phago.RDS"))
#write_csv(dt_summary,here("output","phagocytosis_summary.csv"))
```


```{r include = F}
ggsave(here("manuscript","figures","phago_dunnett_set2.jpg"), dpi = 300, width = 10, height = 8)
```



### Stats - Dunnett's test (non-normalize to protein)

```{r echo = F}
datatable(as.data.frame(emm_contrasts), colnames = c("Contrast", "Mean Diff", "SE", "DF"," Lower CI 95%", "Upper CI 95%","T.ratio","P.adjusted","P.rounded","Sig Level"))

datatable(dt_summary, colnames = c("Treatment","Mean","Standard Deviation (SD)")) %>%
        formatRound(columns = c("mean","sd"), digits = 2)
```


## Phagocytosis Experiment SET 3 {.tabset .tabset-fade .tabset-pills}


### Figure (non-normalize to protein)

```{r echo = F, warning = F, fig.height = 10, fig.width = 12}
# ANOVA
anova <- aov(log2_fluorescence ~ Treatment, data = phago.assay_reorder_3)

# Extract P-value
p_value <- signif(summary(anova)[[1]][["Pr(>F)"]][1],3)

# Contrast with Dunnett
emm_contrasts <- anova %>%
  emmeans("Treatment") %>%
  contrast("trt.vs.ctrl", infer = c(TRUE, TRUE)) %>%
  as_tibble() %>%
  mutate(p.value.round = insight::format_p(p.value, name = NULL,),
         p.value.star = insight::format_p(p.value, name = NULL, stars_only = TRUE)) 

emm_contrasts_split <- emm_contrasts
## Split string
emm_contrasts_split[c('group1', 'group2')] <- str_split_fixed(emm_contrasts_split$contrast, ' - ', 2) 




## Data Summary ##
dt_summary <- phago.assay_reorder_3 %>%
  group_by(Treatment) %>%
        summarize(mean = mean(log2_fluorescence),
                  sd = sd(log2_fluorescence))

################################## Create Plot ##################################

ggplot() + 
        
        geom_bar(data = dt_summary,
    aes(y = mean, x = Treatment, fill = Treatment), stat = "identity", show.legend = FALSE) +
        
        geom_errorbar(data = dt_summary,
                aes(
                  ymin = mean - sd,
                  ymax = mean + sd,
                  x = Treatment
                ),
                width = 0.1) +
        
# Color Bar Plot (Discrete)
  scale_fill_manual(values=as.vector(kelly(20))) + 
        
        labs(x = "Treatment", y = expression(paste("Phagocytosis, ",Log[2],"(Fluorescence)")), subtitle = paste("ANOVA:", " P"," = ", paste0(p_value, collapse = NULL))) +
        
# letters
stat_pvalue_manual(data = emm_contrasts_split,
                   y.position = rep(15,12),
                   label = "p.value.star",
                   x = "group1",
                   vjust = 2,
                   remove.bracket=TRUE,
                   size = 10) +
        theme_classic() +
        theme(axis.text = element_text(angle = 90, color = "black"),
              text = element_text(size = 23)) 
```


```{r include = F}
ggsave(here("figures","phagocytosis_assay","phago_dunnett_set3.jpg"), dpi = 300, width = 10, height = 8)
```



### Stats - Dunnett's test (non-normalize to protein)

```{r echo = F}
datatable(as.data.frame(emm_contrasts), colnames = c("Contrast", "Mean Diff", "SE", "DF"," Lower CI 95%", "Upper CI 95%","T.ratio","P.adjusted","P.rounded","Sig Level"))

datatable(dt_summary, colnames = c("Treatment","Mean","Standard Deviation (SD)")) %>%
        formatRound(columns = c("mean","sd"), digits = 2)
```
