---
title: "Phagocytosis QC"
author: "Brian"
date: "2022-10-18"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 2
---

```{r include = F}
## Packages
if(!any(rownames(installed.packages()) == "pacman")){
   install.packages("pacman")
}

pacman::p_load(here,
               tidyverse,
               DT,
               ggpubr,
               rstatix)
```



```{r include = F}
## Read csv
phago.assay <- read_csv(here("tidy","HMC3 Phagocytosis Compiled.csv"),
                        skip = 2)[,1:4]

phago.assay$Set <- as.factor(phago.assay$Set)
```


## Samples
```{r echo = F}
phago.assay <- phago.assay %>%
        mutate(norm_flu = `Raw Recorded Fluorescence`/Protein_ug_mL)

datatable(phago.assay, colnames = c("Treatment", "Set","RAW Fluorescence","Protein (ug/mL)","Norm Fluorescence per protein")) %>%
        formatRound(columns = c("Protein_ug_mL","Raw Recorded Fluorescence","norm_flu"), digits = 3)
```



## Data Summary 

### Non-normalize to protein
```{r echo = F}
phago.assay.sum <- phago.assay %>%
        group_by(Treatment, Set) %>%
        summarize(mean_flu = mean(`Raw Recorded Fluorescence`),
                  sd_flu= sd(`Raw Recorded Fluorescence`),
                  cv_flu = sd_flu/mean_flu * 100)

datatable(phago.assay.sum, colnames = c("Treatment","Set", "Raw Fluorescence","SD","CV (%)")) %>%
        formatRound(columns = c("mean_flu", "sd_flu","cv_flu"), digits = 3)
```

> CVs are high in non-normalized data. 

***

### Log 2 transform Non-normalize to protein
```{r echo = F}
phago.assay.sum.log <- phago.assay %>%
        group_by(Treatment, Set) %>%
        summarize(mean_flu = mean(log2(`Raw Recorded Fluorescence`)),
                  sd_flu= sd(log2(`Raw Recorded Fluorescence`)),
                  cv_flu = sd_flu/mean_flu * 100)

datatable(phago.assay.sum.log, colnames = c("Treatment","Set", "log2(Fluorescence)","SD","CV (%)")) %>%
        formatRound(columns = c("mean_flu", "sd_flu","cv_flu"), digits = 3)
```

> CVs are much lower in log2(fluorescence) non-normalized data. 

***

## Compare by Sets (non-normalized)

```{r}
ggbarplot(phago.assay.sum, x = "Set", y = "mean_flu",
  label = TRUE, label.pos = "out") +
        labs(y = "Total Fluorescence in Assay")
```

### Mean CVs by set (Log2 non-normalize data)
```{r}
Set_levels <- levels(phago.assay.sum.log$Set)
for (x in Set_levels) {
        hist(phago.assay.sum.log$cv_flu[phago.assay.sum.log$Set==x],
             xlab = paste("Set = ",x),
             main  = paste("CV Histogram: Set = ",x),
             breaks = 10,
             col="darkmagenta")
}

```


### Normalize to protein
```{r echo = F}
norm_phago.assay.sum <- phago.assay %>%
        group_by(Treatment, Set) %>%
        summarize(norm_mean_flu = mean(norm_flu),
                  norm_sd_flu = sd(norm_flu),
                  norm_cv_flu = norm_sd_flu/norm_mean_flu * 100)

datatable(norm_phago.assay.sum, colnames = c("Treatment","Set","Norm  Fluorescence","SD","CV (%)")) %>%
        formatRound(columns = c("norm_mean_flu", "norm_sd_flu","norm_cv_flu"), digits = 3)
```

> Slighlty more variable with protein normalization. 

## Compare by Sets (Normalized)

```{r}
ggbarplot(norm_phago.assay.sum, x = "Set", y = "norm_mean_flu",
  label = TRUE, label.pos = "out") +
        labs(y = "Total Fluorescence in Assay")
```

### Mean CVs by set (Normalize)

```{r}
Set_levels <- levels(norm_phago.assay.sum$Set)

for (x in Set_levels) {
        hist(norm_phago.assay.sum$norm_cv_flu[norm_phago.assay.sum$Set==x],
             xlab = paste("Set = ",x),
             main  = paste("CV Histogram: Set = ",x),
             breaks = 10,
             col="darkmagenta")
}

```

## Summary

> I recommend not combining the sets. Set 2 has lower CVs (and improved after log transformation).

## Save rdata
```{r}
save(phago.assay, phago.assay.sum, norm_phago.assay.sum, file = here("rdata","phago_tidy.rdata"))
```
